---
title: "Select CSAFE Documents"
output: html_document
date: '2022-11-08'
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Overview

This tutorial shows how to use the `select_CSAFE_docs` function to build the 3
sets of documents needed to create a new cluster template, fit a hierarchical
model, and test the model. This function selects documents from the CSAFE
Handwriting Database for the 3 sets - template training documents; model
training documents; and questioned (test) documents. 

The CSAFE Handwriting Database contains 200+ writers. Each writer
wrote three prompts: the London letter, a common phrase, and an excerpt from the
Wizard of Oz. Each writer wrote each prompt three times during three different
sessions. Each session was at least three weeks apart.
The `select_csafe_docs` function allows the user to select the number of writers
to use for the template and model, and ensures that no writers are used in both
sets. The model implemented in handwriter is only applicable to closed-sets of
writers: the questioned document is known to have been written by one writers
whose known handwriting samples are used to fit the model. The
`select_csafe_docs` function selects handwriting samples from the model writers
for use as questioned documents. While the writers are the same, the a document
cannot be used in both the model and the questioned sets. If a user attempts to
do this, `select_csafe_docs` will return an error message.
The `select_csafe_docs` function also allows the user to choose which prompt(s),
session(s), and repetition(s) to use for each of the three sets. As already
mentioned, the function won't allow the user to select the same documents for
the model and questioned sets.

```{r}
docs <- select_csafe_docs(num_template_writers = 10,
                          template_sessions = 1,
                          template_reps = 1,
                          template_prompts = "London Letter",
                          template_seed = 100,
                          num_model_writers = 5,
                          model_sessions = 1,
                          model_reps = c(1,2,3),
                          model_prompts = "Wizard of Oz",
                          model_seed = 101,
                          questioned_sessions = 3,
                          questioned_reps = 1,
                          questioned_prompts = "Wizard of Oz")
```

## Copy sets to a new location
In my workflow, I often want copies of the template training, model training, and questioned docs in a new folder.

```{r}
# folder containing downloaded CSAFE Handwriting database
input_dir <- "~/Documents/non_version_control/CSAFE/data_portal/downloaded"

# output folders
template_dir <- "~/Documents/non_version_control/CSAFE/test_template1/data/template_docs"
model_dir <- "~/Documents/non_version_control/CSAFE/test_template1/data/model_docs"
questioned_dir <- "~/Documents/non_version_control/CSAFE/test_template1/data/questioned_docs"

copy_CSAFE_docs(docs, input_dir, template_dir, model_dir, questioned_dir)
```


